% Plug-in Decomposition

Photran is actually a collection of several Eclipse plug-ins.
The following projects comprise Photran as it is distributed to users.

\begin{itemize}
\item{org.eclipse.photran.core

	This is the Photran Core plug-in.  It
	contains all of the ``behind the scenes'' functionality that
	supports Fortran-specific things in the user interface:
\begin{itemize}
	\item{IAdditionalLanguage for Fortran (i.e., the means
              of adding Fortran to the list of languages recognized
              by the CDT)}
	\item{Error parsers for Fortran compilers}
	\item{Fortran model builder and model elements}
	\item{Interface to Fortran parser (stored in f95parser.jar)}
	\item{Symbol table}
	\item{Module database (and Berkeley DB Java Edition)}
	\item{Refactorings, source printing, program manipulation}
	\item{Non-UI (parsing and model building) preferences}
\end{itemize}}

\item{org.eclipse.photran.managedbuilder.core, org.eclipse.photran.managedbuilder.ui

	This is the Managed Build system.
	Craig Rasmussen (LANL) is handling it.}

\item{org.eclipse.photran.modelicons

	This contains icons for all of the \texttt{FortranElement}s, i.e.,
	the icons used in the Outline and Make Projects views, and
	provides a method which returns \texttt{ImageDescriptor}s for them.}

\item{org.eclipse.photran.ui

	This contains the Fortran-specific components of the user interface:
\begin{itemize}
	\item{Editor}
	\item{Preference pages}
\end{itemize}}

\item{org.eclipse.photran-feature

	This is the Eclipse feature for Photran, used to build the Zip
	file distributed to users.  (A feature is a grouping of related
	plug-ins; in our case, all of the plug-ins listed above.)}

\end{itemize}

\noindent The following projects are in CVS but are not included directly in the
Photran distribution:

\begin{itemize}
\item{org.eclipse.photran-cdt-patches

	This contains CDT language extensibility/neutrality patches.}

\item{org.eclipse.photran-dev-docs

	This contains developer documentation, including this manual.}

\item{org.eclipse.photran-parser

	This contains the Fortran 95 parser, which is exported to
	f95parser.jar and stored in the Core plug-in.
	Note that this plug-in is not included directly in Photran
	and is not (should not be) referenced by any other plug-ins,
	since the parser is referenced through the JAR file
	(see parser description below).}

\item{org.eclipse.photran.tests

	This contains JUnit tests for the parser, symbol table builder,
	source printer, program editor, refactorings, etc.}

\item{org.eclipse.photran.xyzsamplelang

	This was my proof-of-concept for adding an additional language to
	the CDT.  It is completely independent of Photran.  It does
        not exist in the new Photran CVS repository and will soon be
        superceded by the Eightbol project.}

\end{itemize}
